---
title: "Homework 1: Gapminder"
author: "Ziqi Yang"
date: "Monday, September 14, 2015"
output: html_document
---

#1. Gapminder excerpt  
##Bring rectangular data in:  
```{r}
#datat1<-read.table(file="gapminderDataFiveYear.txt")
data2<-read.delim(file="gapminderDataFiveYear.txt")
```

If we use the functions without specifying arguments other than *file*, we can't the same result. For *read.table()*, there are errors. Because when we look into the documents for *read.table()* and *read.delim()*, their arguments are different. The default argument *sep* for *read.table()* is white space, but default argument *sep* for read.delim() is "\\t", which means tab. Because for observation 145, the country's name has space in it, so if we set *sep* argument to white space, R will separate the country and that lead to the error. The second error is for argument *header*, we should set "header=T" in order to exclude the header of the data  out of the observation.  
  
    
This time we can set the argument of *read.table()*, so that it is  equivalent to *read.delim()*:  

```{r}
data1<-read.table(file="gapminderDataFiveYear.txt",sep="\t",header=T)
```

##Small test the data:  
* It is a data.frame. We can see this by "class(data1)"
* Its class is data.frame.
* There are six variables, which are country, year, pop, continent, lifeExp, gdpPercap.
* There are 1698 observations.
* 
  
  
##Explore individual variables:  


